Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-11-18 appendix.tex
tex
sup

In this supplementary material, we provide further technical details and
discussions that are not included in the manuscript due to space limit.
%
Discussions required by the author guidelines can also be found in this
supplementary material.

@@

Adversarial attacks may raise security considerations on
deep learning applications, and even cause negative effect.
%
In contrast, in this manuscript, we focus on defending against the attacks and
enhancing a model's adversarial robustness, and hence is expected to have a
positive impact instead.

~\newline
\noindent\textbf{II. Environment.}

Although adversarial training is not
environment-friendly due to its high training cost (in terms of more intensive
computation than regular training), our method has significantly improved the
adversarial training efficiency compared to the previous works (hence is a
little bit more environment-friendly).

@@

(1) \ul{Triplet Assumption.}

This paper assums training with sample triplets.
%
When other deep metric learning loss functions are used, our proposed method
may not be compatible.
%
This assumption is a limitation of our method.
%
Adversarial training with other DML loss functions is unexplored and is left
for future study.

@@

(1) \ul{Backbone deep neural network model.}

Adversarial training requires more time and GPU memory than regular training.
%
The adversarial training of ResNet-18 with our method requires about 7\~{}10GB
of memory per GPU in distributed data parallel mode.
%
%Adversarial training of ResNet-50 requires much more than 12GB of memory per
%GPU, which will make it harder for some readers to reproduce the results
%and compare.

(2) \ul{Adversarial training budget $\eta$.}

As demonstrated in our experiments, the adversarial training budget, namely the
maximum number of PGD iterations $\eta$ for creating the adversarial examples
can greatly influence the performance.
%
We conduct experiments with multiple choices of $\eta$, and we can see from the
curves that the model performance has a converging trend with large $\eta$
settings.

(3) \ul{Source hardness $H_\mathsf{S}$ and destination hardness
$H_\mathsf{D}$.}

The source hardness $H_\mathsf{S}$ depends on triplet sampling strategy, which
is another research topic in the DML field.
%
The available triplet sampling strategies includes Random, Semihard, Softhard,
Distance-weighted, as well as within-batch Hardest sampling strategies.
%
We conduct experiments and figure out an appropriate choice in our experiments.

The range of choices for $H_\mathsf{D}$ is wider than that of $H_\mathsf{S}$.
%
When we use another benign sample triplet to calculate $H_\mathsf{D}$, the
``another benign sample triplet'' can be sampled with the existing triplet
sampling strategies as well.
%
We tested every combination of sampling strategies in our experiments.
%
Besides, $H_\mathsf{D}$ can be a pseudo-hardness function in order to avoid
disadvantages (discussed in our approach) of using another benign triplet as
reference.
%
Graduate Adversary is a family of such pseudo-hardness functions, and Linear
Graduate Adversary (LGA) is a simple and straightforward example which is
already effective.
%
Although $H_\mathsf{D}$ can be a constant as well, such choice lacks
flexibility and is only used for demonstrating the advantage of LGA in our
experiments.

The $H_\mathsf{S}$ and $H_\mathsf{D}$ are the only two adjustable items of
Hardness Manipulation.

(4) \ul{Parameters in $g_\mathsf{LGA}$ and $g_\mathsf{B}$.}

A parameter $\xi$ is used by $g_\mathsf{B}$ to slightly boost a given
$H_\mathsf{D}$ calculated from another benign triplet.
%
As mentioned in the experiments section, $\xi$ is set to $0.1$ in our
experiment.
%
We refrain from conducting further parameter search because $g_\mathsf{B}$
is only proposed to demonstrate that the problem in the late phase of
training can be alleviated in our step-by-step description of our approach.
%
And $g_\mathsf{B}$ does not alleviate the problem in the early phase of
training, and hence it does not belong to Graduate Adversary.
%
The function $g_\mathsf{B}$ is not a key point.

~\newline
\noindent\textbf{III. Use of Existing Assets.}

@@

@@

\end{comment}

>>>>

In this supplementary material, we provide further technical details and the
less important discussions that are not included in the manuscript due to space
limit.
%
Discussions required by the author guidelines are also included in this
supplementary material.

@@

Adversarial defenses alleviate the negative societal impact of adversarial
attacks, and hence have positive societal impact.

@@

~\newline
\noindent \ul{(1) Triplet Training Assumption.}

Our method assumes sample triplets are used for training.
%
Our method may not be compatible to other non-triplet DML loss functions.
%
Adversarial training with other DML loss functions is left for future study.

~\newline
\noindent \ul{(2) Embedding Space Assumption.}

We follow the common setups~\cite{revisiting,robrank} on the embedding space.
%
Namely, (1) the embedding vectors are normalized onto the real unit
hypersphere;
%
(2) the distance function $d(\cdot,\cdot)$ is Euclidean distance.
%
Our formulations are developed upon the two assumptions.
%
It is unknown whether our method method will be effective when embedding
vectors are \emph{not} normalized.
%
And it is unknown whether our method will be effective when $d(\cdot,\cdot)$ is
replaced as other distance metrics, \eg, cosine distance.

~\newline
\noindent \ul{(1) Maximum number of PGD iterations $\eta$.}

Our method's sensitivity to $\eta$ has been demonstrated by Tab.~3-6 and
Fig.~6-8.
%
A larger $\eta$ indicates higher training cost, and stronger adversarial
examples are created for adversarial training.
%
As a result, a larger $\eta$ leads to a higher robustness (ERS) and a lower
R@1 performance.
%
Our method consistently achieves a higher ERS under different $\eta$ settings
compared to previous methods, and hence are the most efficient defense method.
%
Experiments with $\eta$ larger than $32$ are not necessary because ERS plateaus
according to Fig.~6-8.

~\newline
\noindent \ul{(2) Source hardness $H_\mathsf{S}$ and destination hardness
$H_\mathsf{D}$.}

The $H_\mathsf{S}$ and $H_\mathsf{D}$ are the only two adjustable items in HM.

The source hardness $H_\mathsf{S}$ depends on triplet sampling strategy.
%
We conduct experiments with existing triplet sampling strategies in order to
focus on defense.

The choices for $H_\mathsf{D}$ are more flexible than those of $H_\mathsf{S}$,
as discussed in Sec.~3.1.
%
In the experiments, we study some possible choices following the discussion and
design LGA based on the empirical observations.

~\newline
\noindent\ul{(3) Parameters involved in $g_\mathsf{LGA}$.}

\begin{table*}
\resizebox{\linewidth}{!}{
\setlength{\tabcolsep}{0.47em}
\renewcommand{\arraystretch}{1.1}%
\begin{tabular}{c|cc|cccc|ccccc|ccccc|c}
\hline
\multirow{2}{*}{\textbf{Dataset}} & \multirow{2}{*}{\textbf{Defense}} & \multirow{2}{*}{$\eta$} & \multicolumn{4}{c|}{\textbf{Benign Example}} & \multicolumn{10}{c|}{\textbf{White-Box Attacks for Robustness Evaluation}} & \multirow{2}{*}{\textbf{ERS$\uparrow$}}\tabularnewline
\cline{4-17} \cline{5-17} \cline{6-17} \cline{7-17} \cline{8-17} \cline{9-17} \cline{10-17} \cline{11-17} \cline{12-17} \cline{13-17} \cline{14-17} \cline{15-17} \cline{16-17} \cline{17-17}
 &  &  & R@1$\uparrow$ & R@2$\uparrow$ & mAP$\uparrow$ & NMI$\uparrow$ & CA+$\uparrow$ & CA-$\downarrow$ & QA+$\uparrow$ & QA-$\downarrow$ & TMA$\downarrow$ & ES:D$\downarrow$ & ES:R$\uparrow$ & LTM$\uparrow$ & GTM$\uparrow$ & GTT$\uparrow$ & \tabularnewline
\hline
\hline
\multirow{2}{*}{CUB} & HM{[}$\mathcal{S},g_{\mathsf{LGA}}${]} & 8 & 38.0 & 48.3 & 21.8 & 49.3 & 12.7 & 46.4 & 11.6 & 39.9 & 0.567 & 0.783 & 16.8 & 11.9 & 27.9 & 1.4 & 32.4\tabularnewline
 & HM{[}$\mathcal{S},g_{\mathsf{LGA}}${]} ($u=2.2$) &  & 34.8 & 45.5 & 15.2 & 47.1 & 13.4 & 36.0 & 17.2 & 26.1 & 0.934 & 0.244 & 20.1 & 15.9 & 27.3 & 3.8 & 36.1\tabularnewline
\end{tabular}}
\end{table*}

~\newline
\noindent\ul{Backbone deep neural network model.}

Adversarial training requires more time and GPU memory than regular training.
%
The adversarial training of ResNet-18 with our method requires about 7\~{}10GB
of memory per GPU in distributed data parallel mode.

\subsection{Use of Existing Assets}

\oo{FIXME}

@@

@@

~\newline\textbf{III. Min-max Adversarial Training with Triplet Loss.}

TODO

Collapse even with $\eta=2$ and $\alpha=1$.

@@

[TODO LIST]

Before deadline

* Revise figures

* Revise manuscript

After deadline: supplementary

* add discussions and additional experiments
\end{comment}

\oo{explain lower ERS with high eta? overfit.}

\begin{table*}
\resizebox{\linewidth}{!}{
\setlength{\tabcolsep}{0.47em}
\renewcommand{\arraystretch}{1.1}%
\begin{tabular}{c|cc|cccc|ccccc|ccccc|c}
\hline
\multirow{2}{*}{\textbf{Dataset}} & \multirow{2}{*}{\textbf{Defense}} & \multirow{2}{*}{$\eta$} & \multicolumn{4}{c|}{\textbf{Benign Example}} & \multicolumn{10}{c|}{\textbf{White-Box Attacks for Robustness Evaluation}} & \multirow{2}{*}{\textbf{ERS$\uparrow$}}\tabularnewline
\cline{4-17} \cline{5-17} \cline{6-17} \cline{7-17} \cline{8-17} \cline{9-17} \cline{10-17} \cline{11-17} \cline{12-17} \cline{13-17} \cline{14-17} \cline{15-17} \cline{16-17} \cline{17-17}
 &  &  & R@1$\uparrow$ & R@2$\uparrow$ & mAP$\uparrow$ & NMI$\uparrow$ & CA+$\uparrow$ & CA-$\downarrow$ & QA+$\uparrow$ & QA-$\downarrow$ & TMA$\downarrow$ & ES:D$\downarrow$ & ES:R$\uparrow$ & LTM$\uparrow$ & GTM$\uparrow$ & GTT$\uparrow$ & \tabularnewline
\hline
\hline
\multirow{6}{*}{CUB} & HM{[}$\mathcal{R},\mathcal{M}${]} & 8 & 27.0 & 36.0 & 13.2 & 42.5 & 19.4 & 48.0 & 22.2 & 32.0 & 0.535 & 0.867 & 11.6 & 10.4 & 19.3 & 2.9 & 35.1\tabularnewline
 & HM{[}$\mathcal{R},g_{\mathsf{B}}(\mathcal{M})${]} & 8 & 24.9 & 33.4 & 12.0 & 41.7 & 21.6 & 45.1 & 24.9 & 29.1 & 0.515 & 0.780 & 12.3 & 12.1 & 19.7 & 3.7 & 37.6\tabularnewline
 & HM{[}$\mathcal{R},0${]} & 8 & 2.4 & 3.8 & 1.1 & 21.7 & 40.8 & 22.3 & 42.2 & 16.7 & 1.000 & 0.000 & 1.6 & 1.2 & 1.7 & 7.0 & 43.9\tabularnewline
 & HM{[}$\mathcal{R},-\gamma/2${]} & 8 & 25.5 & 34.5 & 12.8 & 41.7 & 21.9 & 41.5 & 23.8 & 27.0 & 0.493 & 0.784 & 13.5 & 11.8 & 20.3 & 3.6 & 38.3\tabularnewline
 & HM{[}$\mathcal{R},-\gamma${]} & 8 & 27.5 & 36.5 & 13.5 & 42.5 & 18.1 & 54.0 & 19.9 & 37.5 & 0.561 & 0.925 & 9.9 & 9.7 & 18.6 & 1.7 & 32.2\tabularnewline
 & HM{[}$\mathcal{R},g_{\mathsf{LGA}}${]} & 8 & 24.8 & 33.9 & 12.2 & 41.6 & 21.4 & 45.0 & 21.7 & 31.3 & 0.452 & 0.846 & 13.2 & 12.0 & 20.9 & 4.6 & 37.3\tabularnewline
\multicolumn{1}{c}{} & HM{[}$\mathcal{S},g_{\mathsf{1/2}}${]} & 8 & 37.7 & 48.5 & 22.0 & 49.2 & 12.6 & 48.6 & 12.3 & 41.3 & 0.562 & 0.825 & 13.5 & 12.9 & 26.9 & 1.8 & 31.7\tabularnewline
\multicolumn{1}{c}{} & HM{[}$\mathcal{S},g_{\mathsf{LGA}}${]} & 8 & 38.0 & 48.3 & 21.8 & 49.3 & 12.7 & 46.4 & 11.6 & 39.9 & 0.567 & 0.783 & 16.8 & 11.9 & 27.9 & 1.4 & 32.4\tabularnewline
\multicolumn{1}{c}{} & HM{[}$\mathcal{S},g_{\mathsf{2}}${]} & 8 & 37.4 & 48.2 & 17.1 & 49.2 & 12.9 & 45.1 & 13.2 & 39.1 & 0.599 & 0.738 & 17.7 & 12.8 & 27.5 & 1.9 & 33.1\tabularnewline
\multicolumn{1}{c}{} & LGA u=2.2 &  & 34.8 & 45.5 & 15.2 & 47.1 & 13.4 & 36.0 & 17.2 & 26.1 & 0.934 & 0.244 & 20.1 & 15.9 & 27.3 & 3.8 & 36.1\tabularnewline
\end{tabular}}
\end{table*}

\begin{table*}
\resizebox{\linewidth}{!}{
\setlength{\tabcolsep}{0.3em}
\renewcommand{\arraystretch}{1.1}%
\begin{tabular}{ccc|cccc|ccccc|ccccc|c}
\hline
\multicolumn{1}{c|}{\multirow{2}{*}{\textbf{Dataset}}} & \multirow{2}{*}{\textbf{Defense}} & \multirow{2}{*}{$\eta$} & \multicolumn{4}{c|}{\textbf{Benign Example}} & \multicolumn{10}{c|}{\textbf{White-Box Attacks for Robustness Evaluation}} & \multirow{2}{*}{\textbf{ERS$\uparrow$}}\tabularnewline
\cline{4-17} \cline{5-17} \cline{6-17} \cline{7-17} \cline{8-17} \cline{9-17} \cline{10-17} \cline{11-17} \cline{12-17} \cline{13-17} \cline{14-17} \cline{15-17} \cline{16-17} \cline{17-17}
 &  &  & R@1$\uparrow$ & R@2$\uparrow$ & mAP$\uparrow$ & NMI$\uparrow$ & CA+$\uparrow$ & CA-$\downarrow$ & QA+$\uparrow$ & QA-$\downarrow$ & TMA$\downarrow$ & ES:D$\downarrow$ & ES:R$\uparrow$ & LTM$\uparrow$ & GTM$\uparrow$ & GTT$\uparrow$ & \tabularnewline
\hline
{*} & HM{[}$\mathcal{S},g_{\mathsf{LGA}}${]} & 8 & 35.5 & 46.3 & 20.3 & 47.5 & 14.6 & 41.1 & 14.3 & 34.5 & 0.530 & 0.739 & 18.1 & 13.1 & 27.9 & 2.1 & 35.3\tabularnewline
 & +0.5i aap & 8 & 34.8 & 45.2 & 19.8 & 47.6 & 15.6 & 35.9 & 18.5 & 28.2 & 0.802 & 0.447 & 19.2 & 13.9 & 26.3 & 2.9 & 36.4\tabularnewline
 & +i aap & 8 & 34.7 & 45.6 & 19.5 & 46.9 & 16.9 & 34.2 & 18.9 & 26.7 & 0.904 & 0.296 & 20.4 & 16.1 & 26.6 & 4.2 & 37.3\tabularnewline
 & +i aap aan){*}1.0 & 8 & 34.2 & 44.7 & 19.3 & 47.9 & 15.2 & 35.2 & 18.0 & 27.1 & 0.902 & 0.301 & 17.9 & 15.0 & 28.5 & 3.0 & 36.3\tabularnewline
 & +i aa\textasciitilde p apn\textasciitilde{} 1.0 & 8 & 37.9 & 49.4 & 22.5 & 49.2 & 12.9 & 45.9 & 15.0 & 38.9 & 0.909 & 0.378 & 14.8 & 11.7 & 25.3 & 0.9 & 31.4\tabularnewline
 & ICS gamma 0.2 &  & 35.8 & 46.6 & 16.4 & 48.3 & 13.2 & 39.9 & 12.7 & 33.3 & 0.775 & 0.507 & 15.9 & 14.9 & 27.2 & 2.7 & 33.6\tabularnewline
\end{tabular}}
\end{table*}

\begin{table*}
\resizebox{\linewidth}{!}{
\setlength{\tabcolsep}{0.3em}
\renewcommand{\arraystretch}{1.1}%
\begin{tabular}{c|cc|cccc|ccccc|ccccc|c}
\hline
\multirow{2}{*}{\textbf{Dataset}} & \multirow{2}{*}{\textbf{Defense}} & \multirow{2}{*}{$\eta$} & \multicolumn{4}{c|}{\textbf{Benign Example}} & \multicolumn{10}{c|}{\textbf{White-Box Attacks for Robustness Evaluation}} & \multirow{2}{*}{\textbf{ERS$\uparrow$}}\tabularnewline
\cline{4-17} \cline{5-17} \cline{6-17} \cline{7-17} \cline{8-17} \cline{9-17} \cline{10-17} \cline{11-17} \cline{12-17} \cline{13-17} \cline{14-17} \cline{15-17} \cline{16-17} \cline{17-17}
 &  &  & R@1$\uparrow$ & R@2$\uparrow$ & mAP$\uparrow$ & NMI$\uparrow$ & CA+$\uparrow$ & CA-$\downarrow$ & QA+$\uparrow$ & QA-$\downarrow$ & TMA$\downarrow$ & ES:D$\downarrow$ & ES:R$\uparrow$ & LTM$\uparrow$ & GTM$\uparrow$ & GTT$\uparrow$ & \tabularnewline
\hline
\hline
\multirow{2}{*}{SOP} & HM{[}$\mathcal{S},g_{\mathsf{LGA}}${]}\&ICS ($\lambda=0.5$) & 8 & 42.4 & 47.1 & 10.2 & 84.2 & 34.9 & 3.8 & 36.7 & 2.3 & 0.879 & 0.093 & 35.3 & 36.5 & 35.2 & 49.1 & 60.1\tabularnewline
\cline{2-18} \cline{3-18} \cline{4-18} \cline{5-18} \cline{6-18} \cline{7-18} \cline{8-18} \cline{9-18} \cline{10-18} \cline{11-18} \cline{12-18} \cline{13-18} \cline{14-18} \cline{15-18} \cline{16-18} \cline{17-18} \cline{18-18}
 & HM{[}$\mathcal{S},g_{\mathsf{LGA}}${]}\&ICS ($\lambda=0.5$) & 32 & 41.5 & 46.1 & 9.9 & 84.1 & 36.1 & 3.1 & 37.6 & 2.1 & 0.873 & 0.086 & 35.7 & 36.8 & 34.7 & 50.2 & 60.8\tabularnewline
\end{tabular}}
\end{table*}
